---
title: "ASSIGNMENT - 3"
author: "Sai Harshitha Akula"
date: "2023-11-06"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r}
#Importing the dataset.
```
```{r}
Univ_Bank_1 <- read.csv("C:/Users/saiha/OneDrive/Documents/R PROGRAMMING/UniversalBank-1.csv")
```
```{r}
#Loading the required packages.
```
```{r}
library("caret")
library("ISLR")
library("ggplot2")
library("class")
library("lattice")
library("reshape2")
library("melt")
```

```{r}
#Transforming to factor variable.
```
```{r}
Univ_Bank_1$Personal.Loan <- as.factor(Univ_Bank_1$Personal.Loan)
Univ_Bank_1$Online <- as.factor(Univ_Bank_1$Online)
Univ_Bank_1$CreditCard <- as.factor(Univ_Bank_1$CreditCard)
```
```{r}
#Checking the summary of the dataset.
```
```{r}
summary(Univ_Bank_1)
```
```{r}
#Splitting the 100% of data into training and testing.
#60% for training and 40% for validation.
```
```{r}
set.seed(23)
```
```{r}
Split_Index <- createDataPartition(Univ_Bank_1$Personal.Loan, p = 0.6, list = FALSE)
Train_data <- Univ_Bank_1[Split_Index,]
Validation_data <- Univ_Bank_1[-Split_Index,]
```

```{r}
dim(Train_data)
dim(Validation_data)
```
```{r}
normalising1 <- preProcess(Train_data[,-c(10,13:14)],method=c("center","scale"))
Train_data1 <-predict(normalising1,Train_data)
Validation_data1 <-predict(normalising1,Validation_data)
```

A).Create a pivot table for the training data with Online as a column variable, CC as a row 
variable, and Loan as a secondary row variable. The values inside the table should convey 
the count. In R use functions melt() and cast(), or function table(). In Python, use panda 
dataframe methods melt() and pivot().

```{r}
PTable_1 <- ftable(Train_data1[,c(14,10,13)])
PTable_1
```
```{r}
#Here 14- CreditCard, 10- Personal.Loan, 13- Online
```

B).Consider the task of classifying a customer who owns a bank credit card and is actively using 
online banking services. Looking at the pivot table, what is the probability that this customer 
will accept the loan offer?

Ans).Utilizing the data from the pivot table, we can calculate the likelihood of the customer accepting the loan offer as 52 / (52 + 503), resulting in a probability of 0.096.


C).Create two separate pivot tables for the training data. One will have Loan (rows) as a 
function of Online (columns) and the other will have Loan (rows) as a function of CC.
```{r}
melt_1 <- melt(Train_data1,id=c("Personal.Loan"),variable="Online")
melt_2 <- melt(Train_data1,id=c("Personal.Loan"), variable="CreditCard")
```
```{r}
cast_1 = dcast(melt_1, Personal.Loan~Online)
cast_2 <- dcast(melt_2, Personal.Loan~CreditCard)
```

D).D. Compute the following quantities [P(A | B) means “the probability ofA given B”]: 
i. P(CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors)
ii. P(Online = 1 | Loan = 1) 
iii. P(Loan = 1) (the proportion of loan acceptors) 
iv. P(CC = 1 | Loan = 0) 
v. P(Online = 1 | Loan = 0)
vi. P(Loan = 0)

```{r}
ftable(Train_data1[,c(10,13)])

```
```{r}
ftable(Train_data1[,c(10,14)])
```
1. P(CC = 1 | Loan = 1) = (92/92+196) = 0.319
2. P(Online = 1 | Loan = 1) = (167/167+121) = 0.579
3. P(Loan = 1) = (288/288+2712) = 0.096
4. P(CC = 1 | Loan = 0) = (812/812+1900) = 0.299
5. P(Online = 1 | Loan = 0) = (1624/ 1624+1088) = 0.598
6. P(Loan = 0) = (2712/ 2712+288) = 0.904

E). Use the quantities computed above to compute the naive Bayes probability P(Loan = 1 | CC 
= 1, Online = 1).
Ans). (0. 319* 0.579* 0.096) / (0.319 * 0.579 * 0.096) + (0.299 * 0.598 * 0.904) = 0.098


F).Compare this value with the one obtained from the pivot table in (B). Which is a more 
accurate estimate?
Ans).In section B, we obtained a probability value of 0.096, and in the preceding question, we calculated a probability value of 0.098. While these values exhibit slight variations, it's important to note that in part B, we considered a more comprehensive set of dependent information. Therefore, we can confidently assert that the value derived in part B is more accurate and specific in representing the underlying data.

G). Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)? 
Run naive Bayes on the data. Examine the model output on training data, and find the entry 
that corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you 
obtained in (E).
```{r}
library("naivebayes")
```
```{r}
naive_b <- naive_bayes(Personal.Loan~Online+CreditCard,data=Train_data1)
naive_b

```
--The customer who accepts the loan, uses a credit card, and engages in online banking yields a probability of 0.096 according to the Naive Bayes Model. This result closely mirrors the value obtained in section E of our analysis.

